---
output: html_document
editor_options: 
  chunk_output_type: inline
execute:
    freeze: false
    cache: false
---

# Effect of leaf senescence on the timing of leaf unfolding

This additional analysis explores the effect of leaf colouring on the timing of leaf unfolding. The analysis additionally includes the DOY of leaf colouring of the previous year as a fixed effect in the best models of leaf unfolding for beech, larch and linden. I compare shifts in model performance (cAIC, RMSE), significance and direction of effects between best models with the original data set (best model), best models with a reduced data set (best model reduced) and best models updated with the DOY of leaf colouring with a reduced data set (best model senescence). The data set is reduced because of missing pairs between the DOY of leaf unfolding and the DOY of leaf colouring of the previous year.

```{r load-libraries, include=FALSE}
location <- paste0(getwd())
source(paste0("../00_helpers/setup.R")) #common libraries and custom functions
source(paste0("../00_helpers/species_and_phases.R"))

lib_senescence <- c(
  "nlme",
  "cAIC4",
  "broom.mixed",
  "performance",
  "ggeffects",
  "ggtext",
  "ggh4x"
  )
package_fun(lib_senescence, "senescence")

theme_nature <- function() {
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 11, vjust = 0.5, hjust = 0.5),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    strip.background = element_rect(fill = NA, colour = NA, size = 1),
    panel.spacing = unit(0, "lines"),
    strip.text = element_markdown(hjust = 0),
    plot.title = element_text(size = 11),
    axis.title.x = element_text(size = 11),
    axis.title.y = element_blank(), # remove axis title to further reduce space between panels
    plot.margin = margin(t = 0, b = 4, r = -5, l = 0),
    panel.border = element_rect(color = "#E8E8E8", fill = NA, linewidth=1.5),
    panel.grid.minor = element_blank()
  )
}

fun_get_caic <- function(x) {
  tryCatch({
    cAIC(x)$caic # Accessing the specific value from the cAIC function
  }, error = function(e) {
    NA # Return NA or some indication of failure for models that cause an error
  })
}

fun_get_log <- function(x) {
  tryCatch({
    cAIC(x)$loglikelihood
  }, error = function(e) {
    NA # Return NA or some indication of failure for models that cause an error
  })
}

# extract coefficients with confidence intervals
safe_tidy <- possibly(
  ~ broom.mixed::tidy(.x, effects="fixed", conf.int = TRUE) %>% mutate(model_name = .y), 
  otherwise = NULL
)

senescense_tab_fun <- function(data){
  tab <- data %>%
    arrange(SPECIES_SHORT, model_label) %>% 
    mutate_if(is.numeric, signif, 5) %>%
    datatable(rownames = FALSE, filter = 'top', options = list(scrollX = TRUE, dom = 'Bfrtip', pageLength = 10))
  tab
}

random_simple <- as.formula(~ 1 | METEO_ID)
autoregression_simple <- corAR1(form = ~ year | METEO_ID)
```

```{r add-leaf-colouring, include=FALSE}
data <- list.files(path = paste0(path_own_data),
                     pattern = "^data_unscaled_.*\\.rds$",
                     full.names = TRUE) %>%
  setNames(nm = sapply(basename(.), function(x) {
    parts <- unlist(strsplit(x, "_"))
    paste(parts[1:2], collapse = "_")
  })) %>%
  lapply(readRDS) %>% 
  bind_rows() %>%
  # filter complete cases without removing leaf colouring lag
  filter(complete.cases(.)) %>%
  mutate(RELIEF = factor(RELIEF),
         SOCIAL_SITUATION = factor(SOCIAL_SITUATION)) %>% 
  rename_fun()
```

## Model fitting

The fixed effects were fitted by deparsing the terms stored in the original model object and expanding them with the first order orthogonal polynomial of the DOY of leaf colouring of the previous year using a string operation. I then used the original and adapted fixed effects and fitted models with a data set that only includes complete pairs of the DOY of leaf unfolding and the DOY of leaf colouring of the previous year (reduced data set).

```{r load-best, include=FALSE}
best <- list.files(path = paste0(path_output_data),
                   pattern = "^best_con",
                   full.names = TRUE) %>%
    setNames(nm = sapply(basename(.), function(x) {
    parts <- unlist(strsplit(x, "_"))
    paste(parts[1:5], collapse = "_")
  })) %>%
  lapply(readRDS)

best_models <- best %>%
  # transform list into data frame
  enframe(name = "model_name", value = "models") %>%
  unnest_longer(models) %>% 
  rename(submodel_name = models_id, model = models) %>%
  # filter best model and remove spruce
  filter(str_detect(submodel_name, "_1"),
         !str_detect(model_name, "PA")) %>%
  mutate(model_name = str_sub(model_name, 1,11),
         SPECIES_SHORT = str_sub(model_name, 10, 11)) %>%
  select(-model_name) %>%
  rename(model_name = submodel_name) %>% 
  # extract fixed effects for to use in senescense models
  mutate(fixed_effects = map(model, ~deparse(.x[["terms"]]) %>% paste(collapse = " ")) %>% as.character()) # extract fixed effects
```

```{r run-models, filename="Pull original fixed effects and add leaf colouring.", include=TRUE, echo=TRUE}
senescence_models <- data %>%
  group_by(SPECIES_SHORT) %>%
  nest() %>%
  filter(!str_detect(SPECIES_SHORT, "PA")) %>% 
  mutate(fixed_effects_org = case_when(
           SPECIES_SHORT == "FS" ~ best_models[["fixed_effects"]][[1]],
           SPECIES_SHORT == "LD" ~ best_models[["fixed_effects"]][[2]],
           SPECIES_SHORT == "TC" ~ best_models[["fixed_effects"]][[3]],
           TRUE ~ NA_character_),
         fixed_effects_sen = paste0(fixed_effects_org, " + LEAF_COLOURING_LAG")) %>%
  pivot_longer(cols = fixed_effects_org:fixed_effects_sen,
               names_to = "model_name",
               values_to = "fixed_effects") %>%
  rowwise() %>%
  mutate(fixed_effects = list(as.formula(fixed_effects))) %>%
  ungroup() %>%
  mutate(model_result = map2(fixed_effects, data, ~ tryCatch(
           eval(bquote(
             lme(fixed = .(as.formula(.x)),
                 random = ~ 1 | METEO_ID,
                 correlation = corAR1(form = ~ year | METEO_ID),
                 data = .y, method = "ML", na.action = na.omit))),
           error = function(e) e))) %>%
  mutate(model = map(model_result, ~ if (inherits(.x, "error")) NULL else .x),
         error_message = map_chr(model_result, ~ if (inherits(.x, "error")) .x$message else NA_character_))
```

```{r extract-results, filename="Extract results.", include = TRUE, echo = TRUE}
fun_process_models <- function(data){
  data %>%
  filter(map_lgl(model, ~ !is.null(.x))) %>% # filter failed models
  mutate(tidy = map2(model, model_name, safe_tidy), # extract coefficients and p-values
         augment = map2(model, model_name, ~{ # extract fitted values, residuals, fixed effects without data that is not contained in the respective model
           terms <- all.vars(formula(.x))[-1]  # Extract terms and remove response variable
           broom.mixed::augment(.x) %>%
           select(all_of(terms), METEO_ID, year, .fitted, .resid, .fixed) %>%
           mutate(model_name = .y)
           }),
         terms = map(model, ~all.vars(formula(.x))[-1]),
         std_random = map(model, ~safely(~ {
           intervals(.x)$reStruct %>%
             unlist() %>%
             as.data.frame() %>%
             rename(VALUE = 1) %>%
             rownames_to_column("PARAMETER")})(.x)$result), # calculate and extract standard deviation of random effects (variance of random intercept)
         res_random = (res_random = map(model, ~.x$sigma)), # extract residual variance of random effects
         phi = map(model, ~{ # extract autoregressive parameter
           ci = intervals(.x)$corStruct
           tibble(phi = as.numeric(ci[1, "est."]),
                  lower = ci[1, "lower"],
                  upper = ci[1, "upper"])}),
         RMSE = map(model, ~sqrt(mean(.x$residuals^2))), # extract RMSE
         AIC = map(model, ~AIC(.x)), # extract AIC
         logLik = map(model, ~as.numeric(logLik(.x))), # extract logLik
         cAIC = map(model, ~fun_get_caic(.x)), # extract cAIC
         clogLik = map(model, ~fun_get_log(.x)) # extract clogLik
         ) %>%
    unnest(c(AIC, logLik, RMSE,
           cAIC, clogLik)) %>% 
    mutate(coef = map(model, ~coef(.x)),
           terms_count = map(coef, ~ncol(.x)-1),
           collinearity = map(model, ~check_collinearity(.x)),
           RMSE = as.numeric(RMSE))
}

fun_process_output <- function(model_data){
  # extract selection parameters
  selection_parameters <- model_data %>% 
    select(-tidy,
      #-augment, 
      -terms) %>% 
    mutate(calls = map(model, ~deparse(.x[["call"]]) %>% paste(collapse = " ")) %>% as.character(), # extract calls
           fixed_effects = map(model, ~deparse(.x[["terms"]]) %>% paste(collapse = " ")) %>% as.character(), # extract fixed effects
           random_effects = map(model, ~.x[["call"]][["random"]]) %>% as.character()) %>% # extract random effects
    select(SPECIES_SHORT, model_name, model_label, RMSE, AIC, logLik, terms_count, calls, fixed_effects, random_effects,
           cAIC, clogLik) %>%
    unnest(c(RMSE, AIC, logLik, terms_count,
             cAIC, clogLik
             ))

  # extract coefficients and p-values
  tidy <- model_data %>% select(tidy, model_label, SPECIES_SHORT) %>% unnest(tidy)
  
  # extract fitted, fixed and residuals
  augment <- model_data %>% select(augment, model_label, SPECIES_SHORT) %>% unnest(augment)
  
  # extract standard deviation of random effects
  std_random <- model_data %>% select(model_name, model_label, std_random, SPECIES_SHORT) %>% unnest(std_random) %>% pivot_wider(names_from = PARAMETER, values_from = VALUE)
  
  res_random <- model_data %>% select(model_name, model_label, res_random, SPECIES_SHORT) %>% mutate(res_random = as.numeric(res_random))
  
  phi <- model_data %>% select(model_name, model_label, phi, SPECIES_SHORT) %>% unnest(phi)
  
  multicol <- model_data  %>% select(model_name, model_label, collinearity, SPECIES_SHORT) %>% unnest(collinearity)

  return(list(selection_parameters = selection_parameters,
              tidy = tidy, augment = augment,
              std_random = std_random, res_random = res_random, phi = phi, multicol = multicol))
}

senescence_models_processed <- senescence_models %>%
  bind_rows(best_models %>% select(-fixed_effects)) %>%
  mutate(model_label = case_when(model_name %like% "best" ~ "Best model",
                                 model_name %like% "org" ~ "Best model reduced",
                                 model_name %like% "sen" ~ "Best model senescence")) %>% 
  fun_process_models() %>% 
  fun_process_output()
```

```{r debug-ld, eval=FALSE}
# the cAIC cannot be extracted for models using a reduced data set AND including relief position or social situation with the DOY of leaf colouring of the previous year.
fixed_ld <- as.formula(senescence_models[["fixed_effects"]][[4]])

data_ld <- data %>% filter(SPECIES_SHORT == "LD")

fixed_ld01 <- as.formula(doy ~ CDD_conservative + poly(AWC, degree = 2) + poly(STAND_HEIGHT, 
    degree = 2) + SOCIAL_SITUATION + CDD_conservative:poly(GDD_conservative, 
    degree = 2) + 1 + poly(GDD_conservative, degree = 2) + poly(LEAF_COLOURING_LAG, 
    2))

fixed_ld10 <- as.formula(doy ~ CDD_conservative + poly(AWC, degree = 2) + poly(STAND_HEIGHT, 
    degree = 2) + RELIEF + CDD_conservative:poly(GDD_conservative, 
    degree = 2) + 1 + poly(GDD_conservative, degree = 2) + poly(LEAF_COLOURING_LAG, 
    2))

fixed_ld00 <- as.formula(doy ~ CDD_conservative + poly(AWC, degree = 2) + poly(STAND_HEIGHT, 
    degree = 2) + CDD_conservative:poly(GDD_conservative, 
    degree = 2) + 1 + poly(GDD_conservative, degree = 2) + poly(LEAF_COLOURING_LAG, 
    2))

fixed_ld11 <- as.formula(doy ~ CDD_conservative*poly(GDD_conservataive, degree=2) + poly(AWC, degree = 2) + poly(STAND_HEIGHT, 
    degree = 2) + RELIEF + SOCIAL_SITUATION + CDD_conservative:poly(GDD_conservative, 
    degree = 2) + 1 + poly(GDD_conservative, degree = 2))

fixed_ld11 <- as.formula(doy ~ CDD_conservative*poly(GDD_conservative, degree=2) + poly(AWC, degree = 2) + poly(STAND_HEIGHT, 
    degree = 2) + RELIEF + SOCIAL_SITUATION)


model_ld <- lme(fixed_ld11, data=data_ld, correlation = corAR1(form = ~ year | METEO_ID), random = ~ 1 | METEO_ID, method="ML")

cAIC(model_ld)
```

```{r predict-effects}
predictions <- senescence_models %>%
  bind_rows(best_models %>% select(-fixed_effects)) %>%
  mutate(model_label = case_when(model_name %like% "best" ~ "Best model",
                                 model_name %like% "org" ~ "Best model reduced",
                                 model_name %like% "sen" ~ "Best model senescence")) %>% 
  select(-model_result) %>% 
  mutate(terms = map(model, ~ attr(.x$terms, "term.labels"))) %>%
  unnest(terms) %>%
  filter(!terms %like% ":") %>% # remove interaction terms
  mutate(terms = str_remove_all(terms, "poly|\\(|\\)|degree|=|\\d+|\\s|,"),
         prediction = map2(model, terms, ~ tryCatch(
           predict_response(.x, terms = glue::glue("{.y} [all]")) %>% as.data.frame(),
           error = function(e) NULL))) %>%
  select(SPECIES_SHORT, model_name, model_label, prediction, terms) %>% 
  filter(!map_lgl(prediction, is.null)) %>%  # using a regular filter does not work because dplyr returns a single logical value
  mutate(DATA_TYPE = map_lgl(prediction, ~ is.data.frame(.x) && attr(.x, "x.is.factor") == "1")) %>% #identify predictions with categorical variables
  mutate(prediction = if_else(DATA_TYPE, map(prediction, ~ .x %>%
                                               mutate(factor = as.character(x)) %>%
                                               mutate(x = row_number())),
                              prediction)) %>% 
  select(-DATA_TYPE) %>%
  unnest(cols=prediction) %>%
  mutate(Label = label_fun_sim(terms, col = "short"))
```

## Results

```{r tbl-data}
#| tbl-cap: Number ob observations in best models with the original and the reduced data set as well as models fitted with the DOY of leaf colouring.
n_data <- senescence_models %>%
  bind_rows(best_models %>% select(-fixed_effects)) %>%
  mutate(n_data = map_int(model, ~nrow(.x$data)),
         model_label = case_when(model_name %like% "best" ~ "Best model",
                                 model_name %like% "org" ~ "Best model reduced",
                                 model_name %like% "sen" ~ "Best model senescence")) %>%
  select(SPECIES_SHORT, model_label, n_data)

tbl <- n_data %>%
  arrange(SPECIES_SHORT, model_label) %>% 
  datatable(rownames = FALSE)
tbl
```

```{r fig-selection-parameters, fig.height=6, fig.width=6, echo=FALSE, fig.cap = "Comparison of AIC, cAIC and RMSE."}
ggplot(senescence_models_processed[["selection_parameters"]] %>% pivot_longer(cols = c(AIC, cAIC, RMSE), names_to = "PARAMETER", values_to = "VALUE"),
       aes(x = model_label, y = VALUE)) +
  geom_point() +
  facet_wrap(~PARAMETER + SPECIES_SHORT, scales = "free_y") +
  labs(x = "Model name", y = "Value") +
  theme_report(angle = 90)
```

```{r tbl-selection-parameters, echo=FALSE}
#| tbl-cap: RMSE, cAIC and number of terms.
senescense_tab_fun(senescence_models_processed[["selection_parameters"]] %>% select(SPECIES_SHORT, model_label, fixed_effects, everything(), -c(AIC, logLik, calls, random_effects)))
```

```{r fig-coefficients, fig.height= 10, fig.width=8, echo=FALSE, fig.cap = "Comparison of p-values of fixed effects."}
ggplot(senescence_models_processed[["tidy"]], aes(x = model_label, y = term, fill = p.value < 0.1)) +
    geom_tile() +
    geom_tile(color = "#B3B3B3", linewidth = 0.5, fill = NA) +
    scale_fill_manual(values = c("TRUE" = "#A2CD5A", "FALSE" = "#CDC1C5"),
                      labels = c("FALSE" = "not significant", "TRUE" = "significant"),
                      na.value = NA) +
  facet_wrap(~ SPECIES_SHORT)+
    xlab("Model") +
    ylab("Parameter") +
    theme_report(angle = 90, legend.position = "top")
```

```{r tbl-coefficients-fixed}
#| tbl-cap: Fixed effect coefficients and p-values.
senescense_tab_fun(senescence_models_processed[["tidy"]] %>% filter(effect == "fixed") %>% arrange(SPECIES_SHORT, term, model_name) %>% select(SPECIES_SHORT, model_label, term, conf.low, estimate, conf.high, everything(), -c(effect, model_name)))
```

```{r fig-effects, fig.height=8, fig.width=9, fig.cap="Panels show the estimated effect of predictors present in species-specific best models when the DOY of leaf colouring of the previous year is and is not included. Estimates are adjusted for the influence of other included covariates at their mean values. Shaded areas and error bars represent 95% confidence intervals."}
ggplot(predictions %>%
    filter(!is.na(x)), aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high, 
                 color = SPECIES_SHORT, fill = SPECIES_SHORT)) + 
  geom_line(size = 1, aes(group = interaction(SPECIES_SHORT, model_label), linetype = model_label)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, group = interaction(SPECIES_SHORT, model_label)),
              alpha = 0.3, color = NA, fill = "grey") +
  facet_wrap(~ terms, scales = "free_x", strip.position="bottom") +
  labs(x = "", y = "DOY") +
  coord_cartesian(ylim = c(70, 130)) +
  species_color_scale("short")+
  theme_report(legend.position = "top", hjust = 0.5) +
  theme(
  strip.background = element_rect(fill = NA, colour = NA, size = 1),
  strip.placement = "outside",
  panel.spacing = unit(1, "lines"),
  axis.title.x = element_blank(),
  axis.text.x = element_text(margin = margin(t = 3)),
  axis.ticks.length.x = unit(0.2, "cm"),
  panel.border = element_rect(color = "#E8E8E8", fill = NA, linewidth=1.5))
```

```{r fig-effect-senescence, fig.width = 4, fig.height = 4, fig.cap="Estimated effect of leaf colouring in species-specific best models. Estimates are adjusted for the influence of other included covariates at their mean values. Shaded areas and error bars represent 95% confidence intervals."}
ggplot(predictions %>%
    filter(!is.na(x),
           terms %like% "COLOURING"#,
           #SPECIES_SHORT == "FS"
           ), aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high, 
                 color = SPECIES_SHORT, fill = SPECIES_SHORT)) + 
  geom_line(size = 1, aes(group = interaction(SPECIES_SHORT, model_name))) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, group = interaction(SPECIES_SHORT, model_name)),
              alpha = 0.3, color = NA, fill = "grey") +
  labs(x = "Leaf colouring (DOY)", y = "Day of year") +
  coord_cartesian(ylim = c(80, 140))+ # specified as coord carthesian to avoid cutting confidence intervals outside the ylim
  species_color_scale("english")+ 
  theme_report(legend.position = "none", hjust = 0.5) +
  theme_nature()+
  theme(axis.text.y = element_text(size = 10, vjust = 0.5, hjust = 0.5),
        axis.title.y = element_text(size = 11, vjust = 0.5, hjust = 0.5, angle=90, face = "bold"),
        legend.position = "none",
        plot.margin = margin(t = 0, b = 4, r = 7, l = 0),
)
```

```{r fig-effects-size, fig.width=5, fig.height=5, fig.cap="Estimated shift (d) between earliest and latest DOY across the observed range of each environmental variable, based on predictions adjusted to mean covariate values. For each species, colours are scaled so the darkest shade represents the strongest shift."}

effect_size_senescence <- predictions %>%
  filter(model_name == "fixed_effects_sen"#,
         #SPECIES_SHORT == "FS"
         ) %>% 
  group_by(SPECIES_SHORT, terms, Label) %>%
  summarise(MAX_MEAN_EFFECT_SIZE = round(max(predicted) - min(predicted), 2),
            MEAN_STD_ERROR = round(mean(std.error), 2),
            MIN_TERM = x[which.min(predicted)],
            MAX_TERM = x[which.max(predicted)]) %>%
  filter(!is.na(Label),# remove excess labels from categoric variables
         MAX_MEAN_EFFECT_SIZE > 0) %>% 
  rename(PARAMETER = terms) %>% 
  mutate(SPECIES = recode(SPECIES_SHORT, !!!rename_species_en)) %>%
  ungroup() %>% 
  select(Label, SPECIES, everything())

facet_stats <- effect_size_senescence %>%
  group_by(SPECIES) %>%
  summarise(
    min_val = min(MAX_MEAN_EFFECT_SIZE, na.rm = TRUE),
    max_val = max(MAX_MEAN_EFFECT_SIZE, na.rm = TRUE),
    mid_val = mean(MAX_MEAN_EFFECT_SIZE, na.rm = TRUE),
    .groups = "drop")

effect_scaled <- effect_size_senescence %>%
  left_join(facet_stats, by = "SPECIES") %>%
  mutate(scaled_effect = (MAX_MEAN_EFFECT_SIZE - min_val) / (max_val - min_val)) %>%
  inner_join(label_table %>% select(term, param_order_c1, category), by = c("PARAMETER" = "term")) %>% # partial join with order of parameters
  mutate(category = case_when(category == "Soil" ~ "Climate / Soil",
                              category == "Drought" ~ "Climate / Soil",
                              category == "Climate" ~ "Climate / Soil",
                              category == "Phenology" ~ "Pheno",
                              #category == "Stand structure" ~ "Stand",
                              TRUE ~ category),
         Label = case_when(Label %like% "Social" ~ "Social position",
                          TRUE ~ Label),
         order_num = case_when(PARAMETER %like% "ELEVATION" ~ 0, # define parameter order within group
                        PARAMETER %like% "RELIEF" ~ 1,
                        PARAMETER %like% "INCLINATION" ~ 2,
                        PARAMETER %like% "ASPECT" ~ 3,
                        PARAMETER %like% "SOCIAL" ~ 6,
                        PARAMETER %like% "STAND" ~ 7,
                        PARAMETER %like% "HEIGHT" ~ 8,
                        PARAMETER %like% "DROUGHT" ~ 13,
                        PARAMETER %like% "AWC" ~ 12,
                        PARAMETER %like% "CDD" ~ 10,
                        PARAMETER %like% "GDD" ~ 11,
                        TRUE ~ 0),
         PARAMETER = fct_reorder(Label, order_num),
        category = fct_relevel(category, "Pheno", "Stand structure", "Climate / Soil", "Topography")) # order parameter groups

ggplot(effect_scaled, aes(x = SPECIES, y = PARAMETER, fill = scaled_effect)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "#FFA54F", mid = "#FF7F24", high = "#CD661D", midpoint = 0.5, limits = c(0, 1)) +
  geom_text(aes(label = round(MAX_MEAN_EFFECT_SIZE, 2)), size = 4, vjust = 0, colour = "#4F4F4F") +
  xlab("") +
  ylab("") +
  scale_x_discrete(position = "top") +
  facet_nested(rows = vars(category), scales = "free_y", space = "free_y", switch = "y") +
  theme_report(legend.position = "none") +
  theme(
    legend.title = element_text(size = 12, face = "bold", hjust = 0.1),
    axis.text.x = element_text(hjust = 0.5, face = "bold"),
    strip.background = element_blank(),
    strip.text.y.left = element_text(face = "bold", size = 10, hjust = 0.5, colour = "#757575"),
    panel.spacing.y = unit(0.7, "lines"),
    axis.line.y = element_line(colour = "grey", linewidth = 1.2),
    panel.grid = element_blank()
  )
```

1.  The DOY of leaf colouring has a significant delaying effect on the DOY of leaf unfolding of beech, but not larch or linden (@fig-coefficients). If leaf colouring was late in the previous year, leaf unfolding in the following year is also later (@fig-effects, panel leaf colouring lag). The effect is stronger than the effect of AWC (@fig-effects-size). The almost perfect overlap between the same effects of different models, indicated by almost single lines (@fig-effects), indicates that the effect size and shape of other variables is barely affected.
2.  For beech the cAIC is notably better for models including leaf senescense. For linden, the cAIC is worse. For larch the cAIC failed due to the missing data in combination with the factors RELIEF and SOCIAL_SITUATION (@fig-selection-parameters, second row).
3.  The RMSE is close to equal between models with and without leaf senescence and slightly better for models with the reduced data set (@fig-selection-parameters, third row).
4.  The presence of the DOY of leaf colouring does not shift the significance or size of any effects (@fig-coefficients; @fig-effects, see changes in colour between fixed_effects_org and fixed_effects_sen). However, there are shifts in significance between the best model with the original and with the reduced data set for the linear effect of tree height in beech and the quadratic interaction between GDD and CDD for larch (@fig-coefficients). Since the only difference between these two models is the number of observations (@tbl-data), these shifts are likely induced by the missing observations, not the inclusion of leaf senescence.


```{r save-output}
saveRDS(senescence_models_processed[["tidy"]] %>%
          filter(term %like% "COLOURING") %>%
          mutate(label = label_fun_sim(term, "short"),
                 PARAMETER = label) %>% 
          mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
          mutate(p.value = as.character(p.value),
                 p.value = case_when(p.value == 0.000 ~ "≤ 0.001",
                                     TRUE ~ p.value)),
        paste0(path_output_data, "/p_table_senescence.rds"))

saveRDS(effect_size_senescence, paste0(path_own_data, "/effect_size_senescence.rds"))
```
